Figures

### 2016-02-23 doc/paper/results.Rmd

---
title: "Results"
author: "JT McCrone"
date: "Febuary 15, 2016"
output: word_document
---
```{r.message=F,echo=F}
xamp_roc<-"figure 1"
miseq<-"figure 2"
Miseq<-"Figure 2"
hiseq<-"figure 3"
Hiseq<-"Figure 3"
qual<-"figure 4"
Qual<-"Figure 4"
freq<-"figure 5"
Freq<-"Figure 5"
lofreq<-"Figure 6"
Lofreq<-"figure 6"
lower<-"figure 7"
Lower<-"Figure 7"
dups<-"figure 8"
Dups<- "Figure 8"
Table<-"Table 1"
table<-"table 1"
```

Accurately identifying rare single nucleotide variants depends on two components, sensitivity and specificity. Sensitivity, also known as the true positive rate, is the proportion of true variants that are found. Specificity, on the other hand, is the proportion of true negatives that are correctly identified. $1-\text{specificity}$ is the false positive rate, the ratio of false positive variants to true negatives. Both a sensitivity and specificity of 1 describe perfect accurarcy in which all the true variants are found and nothing else.

There is an obvious trade off between sensitivity and specificity. Improved sensitivity often requires less strigent criteria in variant calling which reduces specificity. Conversly, increased stringency can improve specificity but reduce sensitivity. This relationship can be represented as a receiver operating characteristic (ROC) curve (`r xamp_roc`). An ROC plots the sensitivity of an assay along the y axis and 1-specificity along the x axis. To make an ROC curve for a variant callling algorithm one must test the assay on known data. The outcomes can then be stratified according to a metric that quantifies the probability that a given variant is real; often this is a p value or quality score. In a controled benchmakring experiment, all true variants are known, and the sensitivity and specificity can be calculated at different cut offs (`r xamp_roc` A). These points are then used to construct the curve (`r xamp_roc` B). A perfect ROC curve in which all the true positives can be separated from all false positives is a right angle that follows the upper left perimeter of the plot.

@@

It is beyond the scope of this work to comprehensively compare all variant calling methods virologists use. I Instead we focus our efforts towards robustly benchmarking one variant caller and highlighting a few means for improving accuracy. In the process we demonstrate the importance of robustly benchmarking any variant calling pipeline under the experimental conditions to which it is being applied.

 We have chosen DeepSNV as our variant caller of choice for this analysis. DeepSNV uses a clonal, plasmid-derived, control to estimate local error rates controlling for sequence context accross the genome. Because the control is processed together with the experimental samples it has identical noise characteristics. DeepSNV applies a hierarchiacal binomial model at each genomic position and identifies true positive variants with frequencies greater than expected based on the control. Like many variant callers the accuracy of DeepSNV was initially determined using samples that required mimal PCR amplication. To characterize the accuracy of DeepSNV when applied to RT-amplified RNA virus we creadted difed mixutres of two winfluenza strains, WSN33 and PR8 (`r miseq` A).

Because these viruses differ at 491 positions, there are 491 true positives at each dilution. On plasmids subserted to limited amplification, DeepSNV ad a sensitivity 0.860 and a specificity of 1 for SNV at 0.1% frequency. Under our realistic expermental conditions and with a slightly more stringent p value cutoff ( p=0.01 vs. p=0.05) we found a moderate reducttion in sensitivity  (0.830 for variants at 0.63% and 0.14 for variants at 0.16%). Our specificity was above 0.998 in all dilution. While this reduction in specificity appears minor, a false positive rate of 0.002 applied to the xx,xxx potential variants in the 13.xkb influenza genome results in 78 false positives.

@@

Although our first benchmarking experiment validated our ability to accurately detect rare variants in influenza populations, the experiment was run under relatively artificial conditions. The virus populations found in patient-derived samples are expected to be less divergent than the two lab strains. Also the nucleic acid concentration in patient samples is much lower than that found samples grown in cell culture. To mimic these conditions, we generated 20 viral colones, each with a single point mutation in the WSN33 background. We sequenced these mutants on the Illumina Miseq platform and cofirmed that no additional mutations >1% frequency had arisen between transfection and the passage 1 stock. We also quantified the genome copy number of each stock using quantitative PCR. We then mixed equal genome equivalents of these 20 virsuses, generating an artificial population of $10^5$ copies per microliter in which each mutation was present at 5% frequency. This population was serially diluted into a stock of WSN33, generating artificial populations witha ll 20 mutations present a 2,1,0.5,0.2, and 0.1% frequency (`r hiseq` A). We then serially diluted each of these $10^5$ copy number populations into basal media to obtain populations with lower input. The range ($10^3-10^5$ copies per microliter) matches the inpus typically found in pateint-derived samples (data not shown). We sequenced the samples on the Illumina Hiseq platform, and variants were called using DeepSNV. We also processed and sequenced the WSN33 stock in duplicate.

We considered the 20 mutations present in our initial viral mixture as the only true positives in our analysis. Any variant that was present in the wild type stock was masked and neither considered a true positive nor a true negative. This was done to avoid benchmarking DeepSNV with variants identified only be DeepSNV.

The samples with the highest input concentration had the best sensitivity (`r hiseq` B and C). Sensitivity dropped in the 0.5% frequency samples. In general, our sensitivity was lower than that seen previously in the cell culture samples where the input concentration was at least $10^7$ copies per microliter. The drop in sensitivity is likely due to the 100 fold decrease in variant genomes present in the $10^5$ samples. During library prep, a disproportionate amount of the 2.0% sample was lost compared to the others, which likely accounts for the reduced sensitivity relative to the 1.0% sample.

In addition to a more drastic drop in sensitivity we found a much larger number of false positives at a p value cut off of 0.01 (`r hiseq` C). It should be noted that in all cases the specificity was greater than 99%. However, given the large number of potential false positives in an influenza genome (>40,000) even a specificity of 99% results in well over 200 erroneous variant calls. The drop in specificity is likely due to the decease in input levels. In our experience samples grown in cell culture have at least $10^7$ genomes/$\mu$l. As the genomic input decreases, we are more and more reliant on RT-PCR to amplify the true positive "signal" to a detectable level. It is not surprising that is error prone method also amplifies the noise in the process. Initially it is tempting to increase specificity by simply applying a more stringent p value cut off. This would result in marching backwards towards the y axis along the ROCs in `r hiseq` B. However, it is apparent from  `r hiseq` B that any increase in specificity would result in a drastic decrease in sensitivity. It is clear from `r hiseq` that under experimental conditions the DeepSNV output alone is insufficient to accurately separate true positives from false positives.

In deep sequencing studies of viral diversity mapping quality and base quality cutoffs are often used to ensure that only the highest caliber sequencing data is used to call variants. Mapping quality refers to our confidence that a given read is mapped to the correct position in the genome while base quality estimates the likelihood that the base read by the sequencer is correct. In our initial analysis we masked bases that had a base quality score less than 30. We did not apply any mapping quality cut offs. `r Qual` A shows the distribution of the mean quality scores of the variants (p <0.01) present in the $10^5$ genomes/$\mu$l samples. In our case we find that while true and false positive variants can largely be separated by these metrics,  commonly used cut offs are not sufficient (dashed lines `r qual` A). Additionally, we found that in general false positives were biased near the ends of sequencing reads (`r qual` B). This is not surprising as it is well known that sequencing quality dips near the end of a read; however, recall that all bases with phred scores less than 30 were masked from our variant calling. In this case it seems that even high quality false positives are disproportionately called near the ends of reads, while true positives are called uniformly across the read resulting in an average read position near the middle of the read. (_do I need potential reasons for the bias? locations in the genome...)_

In the light of these quality distributions we applied a number of cut offs and drastically increased our specificity without sacrificing sensitivity (`r qual` C and D). For a variant to be considered in our analysis we required a mean mapping quality of 30, a mean phred score of 35, and an average read position within the middle 50% of the read length. _These cut offs were choosen as they ..._ By applying these quality controls we were able to greatly decrease the number of false positives from over 200 in many cases to an average of 9 without sacrificing a single true positive. We recognize that by applying cut offs after variants have been called we are treating potential variant bases more stringently than those matching the consensus sequence. However,because we are applying cut offs to mean quality scores and there are many more consensus bases than variant bases at each position, it is unlikely these biases have any noticable affect on our analysis. In addition to accurately identifying the presence of expected variants DeepSNV was able to  accurately measure the frequency of these variants (`r freq`). However, while the median frequencies are near the expected values, there is a fair spread about each median, something that should be kept in mind when basing down stream analysis on frequency measurements.

#### Lofreq

In addition to DeepSNV, we analyzed our $10^5$ genomes/$\mu$l influenza populations with the variant caller Lofreq. This was done to ensure that the decreased accuracy that accompanied our realistic experimental conditions was not simply based on the DeepSNV variant calling method. Lofreq had marginally decreased sensitivity compared DeepSNV when applied to the variant frequencies $\geq$ 1.0% but marginally increased sensitivity when applied to variant frequencies $\l$ 1.0% (`r lofreq`). Additionally, Lofreq had specificity near that seen in our cell culture benchmarking experiment, and much better than DeepSNV prior to our quality controls. This is not completely surprising as the Lofreq algorithm takes MapQ and Phred scores into account when calling variants, something DeepSNV does not do. Additionally, Lofreq was not prone to calling false positives near the ends of reads, _(verify whether or not it uses read position Show distributions of Lofreq variants?)_, but   `r lofreq` A suggests that sacrificing sensitivity is the only means of increasing specificity.

Lofreq has been reported to have perfect specificity and a sensitivity of 96% for variants present at 0.2%. However, this accuracy reflects a best case scenario using simulated sequencing reads. As was the case with DeepSNV, the accuracy of Lofreq was much lower when applied under conditions that often face virologist using patient derived samples. Our data suggests that is inappropriate to apply Lofreq or DeepSNV under diverse experimental conditions and expect the performance to be unaffected.

It may be possible to further filter the putative variants called by Lofreq and achieve a similar accuracy to what was achieved using DeepSNV. However, because DeepSNV was much less computationally intensive, more user friendly _not sure if I can or should say that_, and capable of high accuracy, we continued to benchmark DeepSNV using the less concentrated samples from our dilution series.

_Include Lofreq in supplimentary?_

#### Lower input levels

Influenza patient samples vary in input concentration by several orders of magnitude and depend on a variety of factors often outside the control of a researcher (day of collection, volume of mucus collected ect.). To ensure accuracy across a range of input levels we diluted our known mutant mixes serially into viral media. We found both decreases in sensitivity and specificity accompanied lower input concentrations (`r lower`). The decreased sensitivity is expected as the likelihood of detecting rare events decreases with sample sizem while the decreased specificity is most likely due to a greater dependence on RT-PCR to amplify the signal to detectable limits. Is it not surprising that an amplification of noise accompanies this dependence. However, our results do highlight the importance of controlling for input levels when comparing populations of RNA viruses.

If RT-PCR errors are responsible for the decreased specificity found at lower input levels, and RT_PCR are randomly distributed across the genome, we would expect such errors to be eliminated by processing the samples in duplicate. To test this hypothesis, a subset of the $10^4$ genomes/$\mu$l were run in duplicate. The 0.2% was not run as we expected a decrease in sensitivity would accompany this approach and had little hope of finding these rare variants based on the sensitivity we saw in `r lower` B. Two RT-PCR reactions were run from a new RNA prep. These samples were processed separately and sequenced on the same lane of an Illumina Hiseq. The previously mentioned quality cut offs were applied and for a variant call to be considered it was required to be found in both duplicates. This drastically improved our specificity and resulted an accuracy that was comparable to that found in the $10^5$ samples (`r dups`). Unexpectedly, an increase in sensitivity accompanied this sequencing run. This is most likely do to a superior library prep with less sample loss during size selection.

>>>>

---
output: pdf_document
---

```{r,message=F,echo=FALSE}
xamp_roc<-"figure 1"
miseq<-"figure 2"
Miseq<-"Figure 2"
hiseq<-"figure 3"
Hiseq<-"Figure 3"
qual<-"figure 4"
Qual<-"Figure 4"
freq<-"figure 5"
Freq<-"Figure 5"
lofreq<-"Figure 6"
Lofreq<-"figure 6"
lower<-"figure 7"
Lower<-"Figure 7"
dups<-"figure 8"
Dups<- "Figure 8"
Table<-"Table 1"
table<-"table 1"

potential_var<-"XX"
```

The ability to reliably distinguish true from false single nucleotide variants (SNV) in next generation sequencing data is integral to accurate studies of viral diversity. Sensitivity is the proportion of true variants identified; while, specificity is the proportion of true negatives identified. $1-\text{specificity}$ is the false positive rate, the ratio of false positive variants to true negatives. Both a sensitivity and specificity of 1 describe perfect accuracy in which all the true variants are found and nothing else.

There is an obvious trade off between sensitivity and specificity. Improved sensitivity often requires less stringent criteria in variant calling which reduces specificity. Conversely, increased stringency can improve specificity but reduce sensitivity. This relationship can be represented as a receiver operating characteristic (ROC) curve (`r xamp_roc`). An ROC plots the sensitivity of an assay along the y axis and 1-specificity along the x axis. To make an ROC curve for a variant calling algorithm one must test the assay on known data. The outcomes can then be stratified according to a metric that quantifies the probability that a given variant is real; often this is a p value or quality score. In a controlled benchmakring experiment, all true variants are known, and the sensitivity and specificity can be calculated at different cut offs (`r xamp_roc` A). These points are then used to construct the curve (`r xamp_roc` B). A perfect ROC curve in which all the true positives can be separated from all false positives is a right angle that follows the upper left perimeter of the plot.

@@

It is beyond the scope of this work to comprehensively compare all variant calling methods employed by virologists. Instead we focus our efforts towards robustly benchmarking one variant caller and highlighting a few means for improving accuracy. In the process we demonstrate the importance of robustly benchmarking any variant calling pipeline under the experimental conditions to which it is being applied.

 We have chosen DeepSNV as our variant caller of choice for this analysis. DeepSNV uses a clonal, plasmid-derived, control to estimate local error rates across the genome. Because it is clonal, the sequence of the control is known with a high degree of confidence and so any nonconsensus base is considered noise. The control is processed together with experimental samples, and so it has identical noise characteristics. DeepSNV applies a hierarchical binomial model at each genomic position and identifies true variants as those with frequencies greater than expected based on the noise found in the plasmid control. Like many variant callers the accuracy of DeepSNV was initially determined using samples that required minimal PCR amplification. We created defined mixtures of two influenza strains, WSN33 and PR8 to characterize the accuracy of DeepSNV when applied to RT-amplified RNA virus populations (`r Miseq` A).

WSN33 and PR8 genomes were mixed serially such that WSN33 was present at relative frequencies of 5, 2.5, 1.25, 0.63 and 0.16%. Because these viruses differ at 491 positions,there are 491 true positives in each dilution. On plasmids subjected to limited amplification, DeepSNV has a sensitivity of 0.860 and perfect specificity for SNV at 0.1% frequency. Under our experimental conditions and with a slightly more stringent p value cutoff ( p=0.01 vs. p=0.05) we found a moderate reduction in sensitivity  (0.830 for variants at 0.63% and 0.14 for variants at 0.16%). The specificity was above 0.998 in all dilutions. While this drop in specificity appears minor, it is important to keep in mind that even a false positive rate of 0.002 results in 78 false positives when applied across the xx,xxx potential variants in the 13.xkb influenza genome.

@@

Although our first benchmarking experiment validated our ability to accurately detect rare variants in influenza populations, the experiment was run under relatively artificial conditions. The virus populations found in patient-derived samples are expected to be less divergent than WSN33 and PR8. Also the nucleic acid concentration in patient samples is much lower than that found samples grown in cell culture. To mimic these conditions, we generated 20 viral colones, each with a single point mutation in the WSN33 background. We sequenced these mutants on the Illumina Miseq platform and confirmed that no additional mutations >1% frequency had arisen between transfection and the passage 1 stock. We also quantified the genome copy number of each stock using quantitative PCR. We then mixed equal genome equivalents of these 20 viruses, to generate an artificial population with $10^5$ copies per microliter and each mutation  present at 5% frequency. This population was serially diluted into a stock of WSN33, generating artificial populations with all 20 mutations present a 2,1,0.5,0.2, and 0.1% frequency (`r Hiseq` A). We then serially diluted each of these populations into basal media to obtain populations with lower input. The range ($10^3-10^5$ copies per microliter) matches the inputs typically found in patient-derived samples (data not shown). We sequenced the samples on the Illumina Hiseq platform, and variants were called using DeepSNV. We also processed and sequenced the WSN33 stock in duplicate to control for any background mutations.

We considered the 20 mutations present in our initial viral mixture as the only true positives in our analysis. Any variant that was present in both duplicates of the wild type stock was masked and  considered neither a true positive nor a true negative. This was done to avoid benchmarking our pipeline with variants identified by our pipeline.

The samples with the greatest input concentration and variant frequencies of 1% and greater had the highest sensitivity (`r hiseq` B and C). Sensitivity dropped at the 0.5% frequency, and in general was lower than that seen previously in the cell culture samples. Our cell culture samples had an  input concentration of at least $10^8$ copies per microliter; whereas, these samples had $10^5$ copies per microliter. The reduced sensitivity is likely due to the fact that there were only 7,000 of each variant genome in the 0.5% frequency samples compared to the equivalent of 62,500,000 in the 0.63% sample from the cell culture experiment.

> This is a little tricky since we amplified the genomes first and then mixed them.

During library prep, a disproportionate amount of the 2.0% sample was lost compared to the others, which likely accounts for the reduced sensitivity relative to the 1.0% sample.

The specificity was greater than 0.99 in all the samples with 10^5^ genomes per microliter. The decreased input concentration and the accompanying increase in RT-PCR cycles is likely the cause of reduced specificity compared with that seen in the cell culture experiment. As above, while 0.99 specificity appears robust, a false positive rate of 0.01 results in over 200 false positive variants when applied to the `r potential_var` potential variants in the influenza genome. Increased specificity can be achieved by applying a more stringent p-value cutoff. However, as shown by the ROC curves in `r hiseq`, this move towards the y axis would markedly reduce sensitivity. Our data demonstrate that with moderate concentrations of input nucleic acid, even statistically significant p-values from a robust variant caller are not sufficient to accurately separate true from false positive variants.

#### Additional filtering criterea

Many next generation sequencing studies utilize mapping (MapQ) and base (Phred) quality cut-offs to ensure that only the highest caliber sequencing data is used to call variants. Mapping quality measures the probability that a given read is mapped to the correct position in the genome while base quality estimates the likelihood that the base call by the sequencer is correct. In our initial analysis we masked bases that had a base quality score less than 30 and did not apply any mapping quality cut offs. We therefore sought to determine if we could improve our specificity by imposing more stringent MapQ and Phred cut-offs. Initially,frequently employed cut-offs such as a MapQ of 20 and a Phred of 30 but were unable to distinguish true from false positives in our 10^5^ samples (`r Qual` A). Many false positives occur on well mapped reads with high quality base calls.

We further parsed our false variant calls by locating them within individual sequencing reads. It is well known that sequence quality drops near the end of a read, and we found that our false positives clustered to these regions (`r Qual` B). The average Phred score of these false positives was xx, and no base calls with a Phred less than 30 was included in the analysis, further demonstrating that filtering on quality score alone is insufficient. In contrast, true positives were uniformly distributed across the reads resulting in an average read position near the middle of the read.

Based on these data we applied a number of empirically determined cut-offs and drastically increased our specificity without sacrificing sensitivity (`r Qual` C and D). For a variant to be considered in our analysis we required a mean mapping quality of 30, a mean phred score of 35, and an average read position within the middle 50% of the read. In addition to these quality cut-offs we applied a number of other criteria in an attempt to increase accuracy such as correcting p-values using the Benjamin Hochman p-value correction, utilizing more stringent p values or frequency cut-offs, retaining PCR duplicate reads, trimming the ends of the influenza genome, and employing various distributions to estimate the error rate in the control sample. All of these results have be summarized in an interactive shiny application available for download at https://github.com/lauringlab/benchmarking_shiny.git and we invite the reader to explore the affects of all these criteria on our data set.

DeepSNV was able to accurately measure the frequency of the true positive variants (`r Freq`). However, while the median frequencies are near the expected values, there is a fair spread in each sample (mean standard deviation of xx). This error should  be kept in mind when employing down stream analysis that depend on frequency measurements such as diversity metrics and comparisons of population structure.

#### Alternative variant callers

DeepSNV is one of many variant callers that use a combination of empiric and statistical approaches to model error rates and separate true from false variants. We asked whether the decreased accuracy observed in our data set was due simply topeculiarities specific to the DeepSNV algorithm. We therefore analyzed our $10^5$ input populations using Lofreq, another alogorthim that is commonly used in viral next generation sequencing studies. Lofreq has been reported to have perfect specificity in the past. Under our conditions Lofreq had marginally reduced sensitivity compared DeepSNV when applied to the variant frequencies \geq 1.0% but marginally increased sensitivity when applied to variant frequencies \l 1.0% (`r Lofreq`). The specificity of LoFreq was comparable to what we observed with DeepSNV in our high-input cell culture-derived populations (`r Miseq`), and better than DeepSNV in our initial implementation (`r Hiseq` prior to Phred, MapQ, and read position filtering). This increased specificity is most likely due to the fast that  the LoFreq algorithm takes MapQ and Phred scores into account when calling variants and has a stringent strand bias filter that removes many of the variants fount only at one end of the reads. However even with these additional characteristics, the specificity of LoFreq was lower than our improved DeepSNV pipeline (`r Qual`) with over 40 false positives per sample. It appears that higher than expected false positive rates are not specific to DeepSNV and most likely plague all variant callers applied to patient-derived viral samples.

\subsection{Lower input levels}

Host-derived viral populations vary in copy number and titer by several orders of magnitude. This variability can be attributed to a variety of factors including but not limited to: collection site, ease of of nucleic acid isolation, library preparation, and host and viral variability. To ensure accuracy across a range of input levels, we diluted our experimental populations serially into basal media and applied our emperically determined quality cut-offs to our list of putative variants (`r Lower`). As expected, the sensitivity for minority variants was lower in populations with lower genome copy number. A variant with a 0.5% frequency in a 10^4^ genomes per microliter sample is only present on 700 genomes and any number of these could be lost in the amplification and library preparation process. We also found that specificity was reduced in samples with lower starting copy numbers. This is presumably due to a greater dependence on RT-PCR amplification. These data highlight the importance of controlling for input levels when comparing diversity across experimental samples.

In most cases, RT-PCRs error should be randomly distributed across the amplified region. If RT-PCR errors are responsible for the reduced specificity found at lower input levels, and RT-PCR are randomly distributed across the genome, we would expect such error to be eliminated by processing the samples in duplicate. To test this hypothesis, we performed duplicate RT-PCR reactions on for the 5,2,1, and 0.5% samples with 10^4^ genomes per microliter. This was done using a separate RNA prep to that shown in `r lower` A. The duplicates were processed separately, but sequenced on the same lane of an Illumina Hiseq. We applied the previously mentioned quality cut-offs and required that a given variant be found in both duplicates. This drastically improved our specificity and resulted in an accuracy that was comparable to that found in the 10^5^ samples (`r Dups`).

### 2016-02-23 scripts/figures.Rmd

---
title: "Figures"
author: "JT McCrone"
date: "January 25, 2015"
output: pdf_document
---

@@

  #plot the roc
  roc.p<-ggplot(roc.df,aes(x=1-specificity,y=sensitivity))+geom_step(color="black",size=1.5)+xlab("1-Specificity")+ylab("Sensitivity")+theme(axis.title.y=element_text(vjust=1))
  roc.p<-roc.p+geom_line(data=lines,aes(x=spec,y=sense,color=as.factor(group)),size=c(1),lty=c(2))+scale_color_brewer(palette="Set1")

@@

```{r,duplicates,fig.height=5,fig.width=3.4}
sum.df<-read.csv("../data/process/2015-11-14/Variants/all.sum.csv",stringsAsFactors = F,comment.char = "#")
#sum.df<-read.csv("~/Box Sync/Lab_work/NGS/Benchmarking/data/2015-11-14/all.sum.csv")
pat<-"([0-9]+)_([0-9]+)"
mutate(sum.df, exp.freq = sub(pat,"\\1",Id),
       dup = as.numeric(sub(pat,"\\2",Id)),
)-> sum.df

sum.df<-subset(sum.df,Id!="PC") # remove the plasmid control that was run
mutate(sum.df,Id=paste(4,exp.freq,sep="_"),exp.freq=ifelse(test = grepl("0",exp.freq),yes=as.numeric(exp.freq)/1000,no = as.numeric(exp.freq)/100))->sum.df

sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants
sum.df$category[sum.df$mutation %in% wt_mut]<-"wt"

cut.df<-subset(sum.df,Phred>35 & MapQ>30 & Read_pos<94 & Read_pos>32)

@@

>>>>

---
title: "Figures"
author: "JT McCrone"
date: "February 11, 2015"
output: pdf_document
---

@@

  #plot the roc
  roc.p<-ggplot(roc.df,aes(x=1-specificity,y=sensitivity))+geom_step(color="black",size=1.5)+xlab("1-Specificity")+ylab("Sensitivity")+theme(axis.title.y=element_text(vjust=1))
  roc.p<-roc.p+geom_line(data=lines,aes(x=spec,y=sense,group=as.factor(group),color=rep(c("#08519cff","#787878ff"),times=3)),size=c(1),lty=c(2))+scale_color_manual(values=c("#08519cff","#787878ff"))

@@

```{r,duplicates,fig.height=5,fig.width=3.4}
dup.df<-read.csv("../data/process/2015-11-14/Variants/all.sum.csv",stringsAsFactors = F,comment.char = "#")
#sum.df<-read.csv("~/Box Sync/Lab_work/NGS/Benchmarking/data/2015-11-14/all.sum.csv")
pat<-"([0-9]+)_([0-9]+)"
mutate(dup.df, exp.freq = sub(pat,"\\1",Id),
       dup = as.numeric(sub(pat,"\\2",Id)),
)-> dup.df

dup.df<-subset(dup.df,Id!="PC") # remove the plasmid control that was run
mutate(dup.df,Id=paste(4,exp.freq,sep="_"),exp.freq=ifelse(test = grepl("0",exp.freq),yes=as.numeric(exp.freq)/1000,no = as.numeric(exp.freq)/100))->dup.df

dup.df<-mutate(dup.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants
dup.df$category[dup.df$mutation %in% wt_mut]<-"wt"

cut.df<-subset(dup.df,Phred>35 & MapQ>30 & Read_pos<94 & Read_pos>32)

@@

### Table 1

Richness
Inverse Simpsons D ( without replacement)
Shannon Entropy
For these I'm assuming all varaints are independent

```{r}
perfect<-data.frame(mutation=rep(true_snv$mutant,times=4),freq.var=rep(c(0.05,0.02,0.01,0.005),each=20))
dups<-subset(dups,p.val<0.01)
dups<-mutate(dups,freq.var=pmax(freq1,freq2))

rich<-function(x){
  dim(x)[1]  
}

H<-function(x){
  p=x$freq.var*log(x$freq.var)
  -sum(p)
}

D<-function(x){
  p=(x$freq.var)^2
  1/sum(p)
}

euc<-function(x,ref){
  x$Id="test"
  ref$Id="ref"
  all<-rbind(subset(x,select=c(mutation,freq.var,Id)),ref)
  tab<-as.matrix(subset(dcast(all,mutation ~ Id, value.var = "freq.var"),select=-c(mutation)))
  tab[is.na(tab)]<-0
  as.dist(dist(tab))
}
get_data<-function(x,setting){
  if(length(unique(x$freq.var))==1){
    exp.freq=unique(x$freq.var)
  }else{
  exp.freq=unique(x$exp.freq)
  }
  out<-data.frame(Richness=rich(x), Entropy=H(x),Simpson=D(x))#,dist=euc(x,subset(perfect,freq.var==exp.freq))[1])
  rownames(out)<-setting
  return(t(out))
}
tabulate<-function(input,freq){
get_data(subset(perfect,freq.var==freq),"Perfect")->perf
get_data(subset(sum.df,gc==input & exp.freq==freq),"DeepSNV")->deepSNV
get_data(subset(lofreq.df,gc==input & exp.freq==freq),"Lofreq")->lo
get_data(subset(cut.read.sig.df,gc==input & exp.freq==freq),"DeepSNV Final")->deepSNV.f
knitr::kable(cbind(perf,deepSNV,lo,deepSNV.f),digits = 2)
}
tabulate.4<-function(input,freq){
get_data(subset(perfect,freq.var==freq),"Perfect")->perf
get_data(subset(sum.df,gc==input & exp.freq==freq),"DeepSNV")->deepSNV
get_data(subset(lofreq.df,gc==input & exp.freq==freq),"Lofreq")->lo
get_data(subset(cut.read.sig.df,gc==input & exp.freq==freq),"DeepSNV Final")->deepSNV.f
get_data(subset(dups,exp.freq==freq),"DeepSNV duplicates")->deepSNV.d

knitr::kable(cbind(perf,deepSNV,lo,deepSNV.f,deepSNV.d),digits = 2)
}

$10^5$ 5%

```{r}
tabulate(5,0.05)
```

$10^5$ 1%

```{r}
tabulate(5,0.01)
```

$10^5$ 0.5%

```{r}
tabulate(5,0.005)
```

$10^4$ 5%

```{r}
tabulate.4(4,0.05)
```

$10^4$ 1%

```{r}
tabulate.4(4,0.01)
```

$10^4$ 0.5%

```{r}
tabulate.4(4,0.005)
```
